---
title: "Exploring Bakeoff Themes With Webscraping"
subtitle: "Webscraping in R Demo"
author: "Martin Ingram"
date: last-modified
date-format: "**DD-MM-YYYY**"
description: "This is a technical demo built using **[Quarto](https://quarto.org/)**"
title-block-banner: "#1C7AE6"
title-block-banner-color: "#FFFFFF"
linkcolor: 'red'
format: 
  html:
    embed-resources: true
    smooth-scroll: true
    code-fold: true
    code-summary: "Reveal code"
    theme: 
      - yeti
    fontcolor: black
    toc: true
    toc-location: left
    toc-expand: 2 
    toc-title: Summary
    toc-depth: 2
---

### Packages

```{r Packages, warning=FALSE}
suppressPackageStartupMessages({
    library(rvest)
    library(httr)
    library(tidyverse)
    # library(tidylog) # Debug only - otherwise gives you loads of filter messages
    library(xml2)
    library(gt)
    library(downloadthis)
    library(reactable)
})

```

# Create links to Bakeoff Season Pages on Wikipedia

Working on the assumption that each URL for each season of Bakeoff follows the same pattern on wikipedia, with a shared root followed by a number:

["https://en.wikipedia.org/wiki/The_Great_British_Bake_Off\_(series_1)"](%22https://en.wikipedia.org/wiki/The_Great_British_Bake_Off_(series_1)%22)

\

Rather than copying each URL we can just append 'n)' to the end of the root for each season we're interested in to build all URLs:

#### **Root:** 

#### "https://en.wikipedia.org/wiki/The_Great_British_Bake_Off\_(series\_"

#### **Season Identifier in URL:**

#### "n)"

```{r Create Links}
# Each season on the wiki page for bakeoff has the same starting structure:  
Bakeoff_Link <- c("https://en.wikipedia.org/wiki/The_Great_British_Bake_Off_(series_")

# To get each link, we only need to paste on the season number to create the full URL:
Link_List = tibble(Season_Number = 1:14, 
                   base_link = rep(Bakeoff_Link, 14), 
                   URL = paste0(base_link, Season_Number, ")")) %>% 
  select(URL)
```

# Scrape Data

Scrape the html data from each URL we build above using the '[rvest](https://rvest.tidyverse.org/)' package, and then clean away noise from our data so we are only left with the following:

-   Season Number

-   Episode Number

-   Theme

```{r Get Data from pages and wrangle into format}
# Container for the episode information: 
Bakeoff_Episodes <- tibble()

for (i in seq_along(Link_List$URL)){
  temp_page <- read_html(GET(Link_List$URL[i], timeout(10))) # read the page

  # process the page:
  temp_page_processed <- temp_page %>%
      html_nodes("a") %>% # The <a> tag (anchor tag) in HTML is used to create a hyperlink on the webpage. 
      html_text("href") %>% # Hypertext reference - The href attribute specifies the URL of the page the link goes to
      tibble(Theme = .) %>% 
      filter(grepl("Episode ", 
                   Theme)) %>% # filter to look for episode titles
      mutate(Season = i, .before = Theme) %>% # add season
      mutate(Episode = sub(x = Theme, 
                           pattern = ":.*",
                           replacement = "")) %>% # get episode number before the colon
      mutate(Episode = sub(x = Episode,
                           pattern = ".* ",
                           replacement = "")) %>% # tidy episode removing excess content
      mutate(Theme = sub(x = Theme,
                         pattern = ".*:",
                         replacement = "")) %>% # Get the theme before the colon
      relocate(Episode, 
               .after = Season) %>% # move the Episode column next to season
      filter(!grepl("[[:digit:]]Episode ",
                    Theme)) %>% # Catch a few errors and remove
      mutate(Theme = str_remove_all(Theme,
    pattern = " \\(Quarter-final\\)|\\(Quarter-Final\\)")) %>% 
      mutate(Theme = str_remove_all(Theme,
    pattern = " \\(Semi-final\\)|\\(Semi-Final\\)")) %>% #Clean themes
      mutate(Theme = str_remove_all(Theme, 
        pattern = " \\(Final\\)"))  %>% 
      mutate(Theme = str_trim(Theme))

  # Add each season to container list of all episodes: 
  Bakeoff_Episodes <- rbind(Bakeoff_Episodes, temp_page_processed)
}
```

# Table of Episodes

```{r show data in table}
# reactable table output
reactable(Bakeoff_Episodes,
          defaultPageSize = 10,
          filterable = TRUE, 
          sortable = TRUE, 
          resizable = TRUE,
          defaultColDef = colDef(align = "right"))
```

```{r Download Button All Data, echo=FALSE}
# option to download data
Bakeoff_Episodes %>% 
  download_this(output_name =
                  paste0(Sys.Date(), " - ",
                  "Bakeoff Data"),
      button_label = "Download Bakeoff Episode Data",
                            output_extension = ".xlsx",
                            button_type = "success",
                            has_icon = TRUE,
                            self_contained = TRUE,
                            icon = "fa fa-save")
```

\

### Make Summaries

```{r make summaries}
# Tidy data and make up the percentage of themes for each episode: 
# i.e. is the the biscuit episode always on the 2nd episode of each season etc.? 
Bakeoff_Episodes <- Bakeoff_Episodes %>% 
  mutate(Theme = recode(Theme, "Cake" = "Cakes")) %>% # harmonise cakes theme
  group_by(Theme, Episode) %>% # make tallies for each theme per episode
  add_tally(name = "Ep_Theme_Count") %>% 
  ungroup() %>%  
  mutate(Total_Eps = max(as.numeric(Season)), # calculate percentage of theme per episode
         "%" = round(Ep_Theme_Count / Total_Eps * 100, 1), 
         Episode = as.numeric(Episode)) %>% 
  group_by(Theme) %>% 
  add_tally(name = "Theme_Raw_Count") %>% 
  ungroup()

# Take unique episode contents for reference
Unique_Theme_Counts_Only <- Bakeoff_Episodes %>% 
  select(Theme, Theme_Raw_Count) %>% 
  distinct() %>% 
  filter(Theme_Raw_Count == 1)

# Group all unique/one-off episode themes together: 
Bakeoff_Episodes <- Bakeoff_Episodes %>% 
  mutate(Theme = if_else(Theme_Raw_Count == 1, "Unique/Others", Theme))

# Get the themes by epsiode data - which % of episodes are each theme? 
Themes_by_Episode <- Bakeoff_Episodes %>% 
  select(Episode, Theme, `%`) %>% 
  distinct()

# Get the themes by season data - what % of themes were in each season?
Themes_by_Season <- Bakeoff_Episodes %>% 
  mutate(Season_Length = if_else(Season == 1, 8, 
                                 if_else(Season == 2, 8, 10))) %>% 
  group_by(Theme, Season, Season_Length) %>% 
  add_tally() %>% 
  ungroup() %>%  
  mutate("%Season" = round(n / Season_Length * 100, 1), 
         Season = as.numeric(Season)) %>% 
  rename("Season_Theme_Count" = n) %>% 
  select(Season, Theme, "%" = "%Season") %>% 
  distinct()

```

\

# Typical Themes Per Episode

```{r Themes by Episode}
# Plot the themes for each episode/ location in the season
plotly::ggplotly(ggplot(Themes_by_Episode, aes(x = Episode, y = `%`)) +
  geom_bar(position="fill", stat="identity", aes(fill = Theme), color = "white") +
  scale_y_continuous(labels = c(0, 25, 50, 75, 100)) +
  scale_x_continuous(breaks = c(1:10), labels = c(1:10)) +
  xlab("Episode") +
  theme_bw() + 
  labs(fill = NULL)+
  scale_fill_viridis_d() +
  labs(title = "Percentage of Themes in Episodes of Bakeoff"),
    tooltip=c("Episode", "Theme"))
```

```{r Download Button Themes_by_Episode, echo=FALSE}
# option to download data
Themes_by_Episode %>% 
  download_this(output_name =
                  paste0(Sys.Date(), " - ",
                  "Bakeoff Episode Data"),
      button_label = "Download Bakeoff Episode Data",
                            output_extension = ".xlsx",
                            button_type = "danger",
                            has_icon = TRUE,
                            self_contained = TRUE,
                            icon = "fa fa-save")
```

\

# Themes In Each Season

```{r Themes by Season}
# Plot the themes for each season: 
plotly::ggplotly(ggplot(Themes_by_Season, aes(x = Season, y = `%`)) +
  geom_bar(position="fill", stat="identity", aes(fill = Theme), color = "white") +
  scale_y_continuous(labels = c(0, 25, 50, 75, 100)) +
  scale_x_continuous(breaks = c(1:13), labels = c(1:13)) +
  xlab("Season") +
  theme_bw() + 
  labs(fill = NULL)+
  viridis::scale_fill_viridis(option="plasma", discrete = TRUE)+
  labs(title = "Percentage of Themes in Episodes of Bakeoff Across Seasons"), tooltip=c("Season", "Theme"))
```

```{r Download Button Themes_by_Season, echo=FALSE}
# option to download data
Themes_by_Season %>% 
  download_this(output_name =
                  paste0(Sys.Date(), " - ",
                  "Bakeoff Season Data"),
      button_label = "Download Bakeoff Season Data",
                            output_extension = ".xlsx",
                            button_type = "info",
                            has_icon = TRUE,
                            self_contained = TRUE,
                            icon = "fa fa-save")

```

```{r Export Data, include=FALSE}
write_csv(Bakeoff_Episodes, paste0(Sys.Date(), " -  Bakeoff Episodes Data.csv"))
```
