---
title: "Time Series Validation"
subtitle: "Prototype Document for validating time series data"
author: "Martin Ingram"
date: last-modified
date-format: "**DD-MM-YYYY**"
description: "This is a technical demo built using **[Quarto](https://quarto.org/)**"
title-block-banner: "#28445e"
title-block-banner-color: "#FFFFFF"
linkcolor: 'red'
format: 
  html:
    embed-resources: true
    smooth-scroll: true
    code-fold: true
    code-summary: "Reveal code"
    theme: 
      - yeti
    fontcolor: black
    toc: true
    toc-location: left
    toc-expand: 2 
    toc-title: Summary
    toc-depth: 2
---

# Organisations Time Series Validation Document

## Notes

-   Useful way of exploring trends at a glance for lots of orgs over time

-   might be susceptible to cases where orgs are missing rows of data for year(s), where sparklines smooth over missing data - this can still be picked up by looking at the datapoints for orgs, but sparklines might be misleading at first glance

    -   Solution, still to test, might be to use complete() on organisation and time to generate NA values for rows/years when data is missing for an org, which forces sparkline to show troughs (for example see Stark Industries trendline which appears smooth but is missing years in the middle)

### Setup Packages

```{r setup packages}
suppressPackageStartupMessages({
  library(tidyverse)
  library(formattable)
 # https://cran.r-project.org/web/packages/sparkline/readme/README.html
  library(sparkline)
  library(DT)
})


options(scipen = 999)
```

### Load Data

```{r load data}
# load data
orgs_data <- 
  read_csv("2023-11-21 - fictional_orgs_time_series.csv", 
           show_col_types = FALSE)

# introduce noise/missing rows
orgs_data <- orgs_data %>% 
  filter(!(organisation %in% c("Abstergo Industries", "LexCorp")
           & time %in% c(2010:2015))) %>% # newly established org
    filter(!(organisation %in% c("Bluth Company", 
                                 "InGen", 
                                 "Pied Piper") & time %in%
               c(2021:2023))) %>% # defunct/closed org
    filter(!(organisation %in%
               c("Florrick Agos and Associates", 
                 "Stark Industries") & time %in%
               c(2016:2018))) %>% # org with missing data 
  mutate(fte = case_when(.default = fte, 
         organisation %in% c("General Hospital", 
                             "Monarch",  # org with incomplete data 
                             "Crichton Davies") &
           time %in% c(2018:2020) ~ NA_integer_)) 
  


# count data availability and calculate volitility
orgs_data <- orgs_data %>% 
    group_by(organisation) %>% 
    arrange(time) %>% 
    mutate(datapoints = n()) %>% 
    mutate(time_range = paste0(min(time), "-", max(time))) %>% 
    mutate(mean_fte = mean(fte,
              na.rm = TRUE), 
           sd_fte = sd(fte, na.rm = TRUE),
    CoV_fte = (sd_fte / mean_fte) * 100) %>%
    mutate(CoV_fte = round(CoV_fte, 1)) %>% 
    group_by(time_range,datapoints, organisation, CoV_fte)

```

## Classifications Breakdown

```{r Show orgs and classifications, results='asis'}

# make up simple table with organisation classifications
orgs_data %>% 
    ungroup() %>% 
    select(organisation, classification) %>% 
    distinct() %>% 
    datatable(escape = F, 
              rownames = F,
              options = list(dom = 'Bfrlipt', 
                             colReorder = TRUE))

```

## Departments

```{r Depts, results='asis'}

# table with sparklines for depts. 
orgs_data %>% 
    filter(classification == "Department") %>%
    summarize(FTE = spk_chr(fte, xvalues = time,
                           tooltipFormat = '{{x}}: {{y}}',
                           lineColor = 'black', 
                           fillColor = 
                               tinter::lighten('black', 
                                               amount = 0.3),
                           chartRangeMin = 0,
                           chartRangeMax = 10,
                           width = 80,
                           height = 60,
                           highlightLineColor = 'orange', 
                           highlightSpotColor = 'orange'),
              .groups = c("keep")) %>% 
    datatable(escape = F, 
              rownames = F,
              options = list(dom = 'flipt', 
                             colReorder = TRUE,
                             filter = 'top',
fnDrawCallback = htmlwidgets::JS('function(){
                         HTMLWidgets.staticRender(); }'))) %>% 
    spk_add_deps()
```

## Agencies

```{r Agencies, results='asis'}

# table with sparklines for Agencies 
orgs_data %>% 
    filter(classification == "Agency") %>% 
    summarize(FTE = spk_chr(fte, xvalues = time,
                           tooltipFormat = '{{x}}: {{y}}',
                           lineColor = 'darkblue', 
                           fillColor = 
                               tinter::lighten('darkblue', 
                                               amount = 0.3),
                           chartRangeMin = 0,
                           chartRangeMax = 10,
                           width = 80,
                           height = 60,
                           highlightLineColor = 'orange', 
                           highlightSpotColor = 'orange'),
              .groups = c("keep")) %>% 
    datatable(escape = F, 
              rownames = F,
              options = list(dom = 'flipt', 
                             colReorder = TRUE,
                             filter = 'top',
fnDrawCallback = htmlwidgets::JS('function(){
                         HTMLWidgets.staticRender(); }'))) %>% 
    spk_add_deps()
```

## External Bodies

```{r External Bodies, results='asis'}

# table with sparklines for external bodies 
orgs_data %>% 
    filter(grepl(classification,
                 pattern = "External Body")) %>% 
    summarize(FTE = spk_chr(fte, xvalues = time,
                           tooltipFormat = '{{x}}: {{y}}',
                           lineColor = 'darkred', 
                           fillColor = 
                               tinter::lighten('darkred', 
                                               amount = 0.3),
                           chartRangeMin = 0,
                           chartRangeMax = 10,
                           width = 80,
                           height = 60,
                           highlightLineColor = 'dodgerblue', 
                           highlightSpotColor = 'dodgerblue'),
              .groups = c("keep")) %>% 
    select(time_range, datapoints, CoV_fte, FTE, organisation) %>% 
    datatable(escape = F, 
              rownames = F,
              options = list(dom = 'flipt', 
                             colReorder = TRUE,
                             filter = 'top',
fnDrawCallback = htmlwidgets::JS('function(){
                         HTMLWidgets.staticRender(); }'))) %>% 
    spk_add_deps()
```

## Organisations in Scope Per Year

```{r Org Counts Per Year}

# table with number of orgs in scope for each year
orgs_data %>% 
    group_by(time) %>% 
    summarise(org_count = n()) %>% 
    select(time, org_count) %>% 
    datatable(escape = F, 
              rownames = F,
              options = list(dom = 'Bfrlipt', 
                             colReorder = TRUE))
```

## Classifications In Scope Per Year

```{r Classifications Per Year}

# table with number of orgs per classification in scope for each year

crosstab_classifications <- orgs_data %>% 
    janitor::tabyl( classification, time)

crosstab_classifications %>% 
    datatable(escape = F, 
              rownames = F,
              options = list(dom = 'Bfrlipt', 
                             colReorder = TRUE))

downloadthis::download_this(crosstab_classifications,
        output_name = paste0(Sys.time(), 
        " - Classification_by_Year_Totals"),
        output_extension = ".xlsx",
        button_label = 
            "Download Classification Totals Data",
         button_type = "success",
        has_icon = TRUE,
        self_contained = TRUE,
        icon = "fa fa-save")
```

## FTE Totals by Classification and Year

```{r Summary FTE Table Data}

# Get totals and counts for all organisations overall
all_orgs_totals <- orgs_data %>% 
  group_by(time) %>% 
  summarise(fte_total = sum(fte, na.rm = TRUE), orgs_by_year = n(), .groups = "keep") %>% 
  mutate(classification = "Total")

# get totals and orgs for each classification and
# combine with totals above
Org_Clas_Totals <- orgs_data %>% 
  group_by(time, classification) %>% 
  summarise(fte_total = sum(fte), orgs_by_year = n()) %>% 
  rbind(all_orgs_totals) %>% 
  mutate(classification = factor(classification, 
                levels = c("Total",
                           "Department", 
                           "Agency", 
                           "External Body"))) %>% 
  arrange(time, classification)

```

```{r Summary FTE Table}
# show summary/totals in table
Org_Clas_Totals %>% 
    datatable(escape = F, 
              rownames = F,
              options = list(dom = 'Bfrlipt', 
                             colReorder = TRUE))

downloadthis::download_this(Org_Clas_Totals,
        output_name = paste0(Sys.time(), 
        " - All_Org_Totals"),
        output_extension = ".xlsx",
        button_label = "Download FTE Total Data",
         button_type = "primary",
        has_icon = TRUE,
        self_contained = TRUE,
        icon = "fa fa-save")
```

## All Data

```{r Show all data, results='asis'}
# Show all data for exploring issues
orgs_data %>% 
    ungroup() %>% 
    select(organisation, 
           time, parent_company, 
           classification, CoV_fte, fte) %>% 
datatable(escape = F, 
              rownames = F,
              options = list(dom = 'Bfrlipt', 
                             colReorder = TRUE))

```

```{=html}
<style>
table.display td { white-space: nowrap; }
</style>
```
