## *`r organisation`*

Stats for **`r organisation`**:

-   Classification: `r classification`

-   CEO: `r data$fake_org_ceo_name`

-   Contact Number: `r data$fake_contact_number`

-   Staff in `r data$time`: `r data$label_fte`

-   Running Costs in `r data$time`: £`r data$label_running_costs`

-   Investment Costs in `r data$time`: £`r data$label_investment_costs`

-   Funding in `r data$time`: £`r data$label_funding`

### `r organisation`: FTE By Running Costs

```{r}
#|: warning: FALSE
# cant give chunks names as they are duplicated by the code
# scatter plot
org_highlight <- organisation
clas_highlight <- classification

plot_color_hightlight <- case_when(.default = "green1",
                                   classification ==
                                       "Agency" ~ "dodgerblue",
                                   classification ==
                                       "Department" ~ "firebrick",
                                   classification ==
                                       "External Body" ~ "seagreen")

ggplot2::ggplot(latest_orgs_data,
                aes(x = fte, y = running_costs, fill = classification))+
    geom_jitter(size = 4.5, shape = 21, fill = plot_color_hightlight)+ 
  scale_x_continuous(labels =
                         label_number(scale_cut = cut_short_scale()))+
  scale_y_continuous(labels =
                         label_number(scale_cut = cut_short_scale()))+
    gghighlight::gghighlight(organisation == org_highlight, label_key = organisation,
                             use_group_by = FALSE, use_direct_label = TRUE)+
    theme_minimal()+
    xlab("Staff Employed (FTE)") +
    ylab("Running Costs (£)")
```

\<br\>

### `r organisation`: FTE Over Time

#### Compared Against Other `r clas_highlight` Organisations

```{r}
#|: warning: FALSE
# time series

orgs_data %>% 
    filter(classification == clas_highlight) %>% 
    ggplot2::ggplot(aes(x = time, y = fte))+
    geom_line(aes(group = organisation), color = plot_color_hightlight, linewidth = 1.4)+
    gghighlight::gghighlight(organisation == org_highlight, label_key = organisation, 
                             use_group_by = FALSE, use_direct_label = FALSE)+
    theme_minimal()+
    xlab("Year (March 31st)") +
    ylab("Staff Employed (FTE)") +
    geom_point(data = latest_orgs_data %>% 
                 filter(classification == clas_highlight, 
                        organisation == org_highlight),
               aes(x = time, y = fte, fill = classification), 
               size = 4.5, shape = 21, fill = plot_color_hightlight)
    # geom_label(data = filter(latest_orgs_data,
    #                          organisation == org_highlight), 
    #            aes(label = organisation), 
    #            fill = plot_color_hightlight, 
    #            color = "white",
    #            nudge_x = -1.4,
    #            nudge_y = -10)
```

\<br\>

```{r}
# waffle plot of funding to costs 
#| fig.align: center
#| fig-height: 5.5
#| fig-width: 6.3

test <- latest_orgs_data %>% 
    filter(organisation == org_highlight) %>% 
    pull(costs_to_funding_proportion) %>% 
    append(100- .)

names(test) <- c("(%) Funded", "(%) Non-Funded")

waffle::waffle(test, rows = 10, flip = TRUE, 
               colors = c(plot_color_hightlight,
                          "grey80"), 
               legend_pos = "bottom") +
  theme(legend.text = element_text(size = 20))
```

\<br\>
